Accessing the nucleon transverse structure in inclusive deep inelastic scattering

### 2017-07-10 1607 - JMC for g2/Paper/JMC_dis.tex
FIxed references, and cleanup

@@

\date{\bf Thursday, May 4th, 2017}
%\date{\bf \today, \currenttime}

\begin{abstract}
We revisit the standard analysis of inclusive Deep Inelastic Scattering
off nucleons taking into  account the fact that on-shell quarks cannot be present in the final state, but they rather decay into hadrons --  a process that can be described in terms of suitable ``jet'' correlators. As a consequence, a spin-flip term
associated with the invariant mass of the produced hadrons is generated
non perturbatively and couples to the target's transversity distribution
function. In inclusive cross sections, this provides an hitherto neglected and
large contribution to the twist-3 part of the $g_2$ structure function, that
can explain the discrepancy between recent calculations and fits of this
quantity.  It also provides an extension of the Burkhardt--Cottingham sum rule, putting constraints on the small-$x$ behavior of the transversity function, as well as an extension of the Efremov--Teryaev--Leader sum rule, suggesting a novel way to measure the tensor charge of the proton.
%Perspectives to measure the new spin-flip term, and applications to
%spin-1 targets will be briefly discussed. 
\end{abstract}

@@

The transversity distribution is notoriously difficult to access because it is
a chiral-odd function and needs to be combined with a spin-flip mechanism to
appear in a scattering process~\cite{Jaffe:1996zw}. Usually, this spin flip is provided by another
nonperturbative distribution or fragmentation function, accessible in Drell-Yan or semi-inclusive Deep Inelastic Scattering (DIS)~\cite{Ralston:1979ys,Jaffe:1991kp,Jaffe:1993xb,Collins:1992kk}.
The only other known way to attain spin-flip terms in Quantum Electro-Dynamics and QCD is taking into account mass corrections. In fact, it is well known that the transversity distribution gives a contribution to the structure function $g_2$ in inclusive DIS (see, \eg, \cite{Accardi:2009au} and references therein), and in particular to the violation of the so-called Wandzura--Wilczek relation for $g_2$~\cite{Wandzura:1977qf}. However, this contribution is proportional to the current quark mass and can be expected to be negligibly small.
%This implies that the transversity distribution cannot be accessed in inclusive DIS, but only in more complex processes such as semi-inclusive DIS or dilepton production in hadron-hadron collisions.

@@

\begin{figure*}[bt]
  \centering
  (a)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram0.eps}
  \hfill
  (b)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram2.eps}
  \hfill
  (c)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram1.eps}
  \caption{Diagrams contributing to inclusive DIS scattering up to twist-3, including a jet correlator in the top part. Note the gluon attaches to both the nucleon and jet correlators. The Hermitian conjugates of diagrams (b) and (c), i.e., with gluons attaching to the right of the cut, are not shown.
  }
  \label{fig:handbags}
\end{figure*}

When inserting the jet correlator in the handbag diagram for inclusive DIS, the integration over $dl^+$, or equivalently $dl^2/(2l^-)$, is kinematically coupled to the other integrations, and induces corrections of order ${\cal O}(1/Q^2)$ whose effect on the $F_2$ structure function has been studied in Ref.~\cite{Accardi:2008ne}. In this paper, where we limit our
attention to effects of order ${\cal O}(1/Q)$, we can neglect $k^-$ compared to $q^-$. As a consequence, we can extend the integration over $dl^2$ to infinity, with the consequence that the jet correlator decouples from the parton correlator $\Phi$, and the inclusive structure functions only depend on the integrated jet correlator 
\begin{equation} 
  \Xi(l^-,{\bm l_T}) \equiv \int \frac{dl^2}{2l^-} \, \Xi(l) 
    =  \frac{\Lambda}{2 l^-}\,\xi_1 \id
    +  \xi_2 \frac{\nslash_-}{2} 
    + {\cal O}(l_T/l^-) + \text{higher\ twists} \ .
\end{equation} 
The neglected ${\bm l_T}$-dependent and higher twist terms only contribute to ${\cal O}(1/Q^2)$ to the inclusive cross section. Note that thanks to Eq.~\eqref{eq:jetfnsprops} we obtain
\begin{align}
\xi_1 &= \int d\mu^2 \frac{\mu}{\Lambda} J_1(\mu^2) 
       \equiv \frac{\mj}{\Lambda},
&
\xi_2 &= \int d\mu^2 J_2(\mu^2) = 1 \ ,
\end{align} 
where $\mj$ can be interpreted as the average invariant mass produced in the spin-flip fragmentation processes of a quark of flavor $q$.
It is important to notice that while $\xi_2=1$ exactly due to CPT invariance
(see Sec.~10.7 of Ref.~\cite{Weinberg:1995mt}), the jet mass $0 < \mj < \int d\mu^2 \mu J_2(\mu^2)$ is dynamically determined. From the analytic properties of the spectral functions we know that $J_2(\mu^2) = Z \delta(\mu^2-m_q) + \bar J_2(\mu^2) \theta (\mu^2-m_\pi^2)$ with $Z<1$ and the continuum starting at $m_\pi$ (the mass of the pion) due to color confinement effects. Taking into account that $J_1 < J_2$, we may therefore estimate 
\begin{align}
  \label{eq:mjet}
  \mj = {\cal O}(100 \text{\ MeV}) \ .
\end{align}
Although $\mj$ is in general a nonperturbative quantity, it is interesting to
notice that on the perturbative vacuum, 
\begin{align*}
  \Xi^\text{pert}(l) = (\lslash + \mq \id) \, \delta(l^2-m_q^2) + {\cal O}(\alpha_s) \ ,
\end{align*}
where $\mq$ is the current quark mass; therefore $\mj^\text{pert}=m_q$, and one recovers the result of the calculation with the conventional handbag diagram.
% \begin{align}
%   \label{eq:xi2_chiral_cond}
%   \mj = \frac{\Lambda}{4} \int \Tr \big[ \Xi(l) \bm 1] 
%    = \langle 0 | \bar \psi_i(0) \psi_i(0) | 0 \rangle
% \end{align}
% % \todo{[AA] 2 issues here: (1) possible factors of 4. 
% % (2) I am neglecting the sum over quarks; we have to fix that somehow but there is a little tension between the quark level $\xi^a$ and the contribution to structure function which is summed and weighted by the electron charge.} 
% On the non-perturbative vacuum, this corresponds the contribution of a quark of flavor $q$ to the chiral condensate, and is calculable in principle in lattice QCD.
% Considering the perturbative vacuum and calculating the same to leading order in the strong coupling constant, corresponds, instead, to taking the trace of the cut bare-quark propagator to obtain 
% $\mj = \,_{\text{pert}} \langle 0
% |  \bar \psi_i(0) \psi_i(0) | 0 \rangle_{\text{pert}}  = \mq$, with $\mq$ the quark mass. One recovers, therefore, the result of the calculation with the conventional handbag diagram. 
However, we are here considering non perturbative effects
in the quark propagation, and $\mj \gg \mq$. 
Therefore, differently from $J_2$, the $J_1$ function leaves an imprint on the inclusive DIS cross section even in the asymptotic $Q^2 \to \infty$ regime.

@@

The role of the $\xi_1=M_q/\Lambda$ term in inclusive DIS can be discussed by analyzing the following terms of the semi-inclusive hadronic tensor:
\begin{align}
  \label{eq:Wsidis_ini}
  2 \Lambda  W^{\mu\nu}
    & = i \frac{2\Lambda}{Q} \hat t^{[\mu}_{\phantom \perp} 
    \epsilon_\perp^{\nu]\rho}S_{\perp\rho} \\
    & \times \sum_q e_q^2
    \bigg[ 2 \xbj g_T^q(\xbj) \sum_h \int dz d^2p_{hT} D_1^{q,h}(z,p_{hT}) 
  %\\ &
  + 2 h_1^q(\xbj) \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) \bigg] + \ldots \ , \nonumber
\end{align}
where $g_T^q(z,p_{hT})$ and $\tilde E^q(z,p_{hT})$ are twist-3 TMDs originating, respectively, from the quark-quark and the quark-gluon-quark fragmentation correlators. 
For clarity, here we reintroduced the quark flavors $q$, $e_q$ being their respective electric charge.
The first term can be easily integrated with the help of the sum rules
\eqref{eq:SIDIS_to_DIS_TMDlevel1} and \eqref{eq:SIDIS_to_DIS_TMDlevel2}. To integrate the second term, however, we first need make
use of the relation $\tilde E(z) = E(z) - (\mq/\Lambda) z D_1(z)$, which is a
consequence of the QCD equations of motion \cite{Bacchetta:2006tn}, then
make again use of the sum rules \eqref{eq:SIDIS_to_DIS_TMDlevel1}-\eqref{eq:SIDIS_to_DIS_TMDlevel2} to obtain
\begin{align}
  \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) 
    = \xi_1 - \frac{\mq}{\Lambda} \xi_2 = \frac{\mj - \mq}{\Lambda }\ .
\end{align}
This formula provides a
non perturbative generalization of the commonly used $\int\tilde E =0$ sum
rule introduced in \cite{Jaffe:1996zw}. Indeed, calculating the jet correlator 
on the perturbative vacuum one would obtain, as already discussed, $\mj=\mq$
and the integral would vanish.

Finally, the contraction of the hadronic tensor with the leptonic tensor leads
to the following result for the inclusive DIS cross section up to order $\Lambda/Q$~\cite{Bacchetta:2006tn}:
\begin{align}
\frac{d\sigma}{d\xbj \, dy\, d\phi_S}
%&
=
\frac{2 \alpha^2}{\xbj y Q^2}\,
\frac{y^2}{2\,(1-\varepsilon)}\, 
\biggl\{
&F_{T} + \varepsilon F_{L}
+ S_\parallel \lambda_e\,
  \sqrt{1-\varepsilon^2}\; 
F_{LL}
%\nonumber 
%\\  &
+ |\bm{S}_\perp| \lambda_e\, \sqrt{2\,\varepsilon (1-\varepsilon)}\, 
  \cos\phi_S\, 
F_{LT}^{\cos \phi_S}
 \biggr\} \ ,
\label{e:crossdis}
\end{align}
where $\phi_S$ is the angle between the transverse component of the proton spin vector and the lepton plane, $\epsilon$ is the ratio of the longitudinal and transverse photon fluxes, and $\lambda_e$ is the electron's helicity. 
The structure functions on the right hand side read
\begin{align}
F_{T} &= \xbj\,\sum_q e_q^2\,f_1^q(\xbj),
\\
F_{L} &= 0,
\\
F_{LL} &=\xbj\,\sum_q e_q^2\,g_1^q(\xbj),
\label{e:FLLint}
\\
F_{UT}^{\sin \phi_S}&=0,
\label{e:FUTint}
\\
F_{LT}^{\cos \phi_S}&=-\xbj\,\sum_q e_q^2\, \frac{2\Lambda}{Q}\,
\biggl(\xbj  g_T^q(\xbj)
   + \frac{\mj -\mq}{\Lambda} \, h_{1}^q(\xbj) \biggr) \ ,
\label{e:FLTint}
\end{align}
where $f_1^q$, $g_1^q$ and $h_1^q$ are the unpolarized, polarized, and transversity PDFs,respectively.
The second term in the last structure function is a new result from our
analysis, it is proportional to the jet mass, and it is not suppressed as an
inverse power of $Q$ compared to the 
standard $g_T$ term. Perturbatively, $M_q^\text{pert}=m_q$ and the new term vanishes. However, on the nonperturbative vacuum the jet mass $M_q$ is much larger than the quark's current mass $m_q$, originating a nonnegligible term to the twist-3 part of the target's $g_2$ structure function, as we will discuss in the next section.

@@

The structure functions in Eqs.~\eqref{e:FLLint} -- \eqref{e:FLTint} can be
related to the usual
structure functions $g_1$ and $g_2$ defined from the following Lorentz decomposition of the hadronic tensor
\begin{align}
\begin{split}
W^{\mu\nu}(P,q) 
  & = \frac{1}{P\cdot q} \varepsilon^{\mu\nu\rho\sigma} q_\rho 
  \Big[ S_\sigma g_1(x_B,Q^2) 
    + \Big( S_\sigma - \frac{S\cdot q}{P\cdot q}\, p_\sigma
	\Big) g_2(x_B,Q^2)
  \Big] \ ,
\label{eq:Wmunu}
\end{split}
\end{align}
Then, in particular,
\begin{align}
  g_1 &= \frac{1}{x_B} F_{LL} \\
  g_2+g_1 &= \frac{Q}{2\Lambda x_B^2} F_{LT}^{\cos \phi_S} \ .
\end{align}
% \begin{align}
% g_1&=\frac12 \sum_q e_q^2g_1^q \ , \\
% g_2&=\frac12 \sum_q e_q^2(g_T^q-g_1^q) \ .
% \end{align}
Utilizing Equations of Motion and Lorentz Invariance relations as discussed in Ref.~\cite{Accardi:2009au}, we arrive at
\begin{align}
\label{e:g2}
  g_2(\xbj) = g_2^{WW}(\xbj) + \frac{1}{2}\,\sum_a e_a^2
\biggl(
    \widetilde g_T^{a \star}(\xbj) 
    + \int_\xbj^1\frac{dy}{y} \widehat{g}_T^q(y) 
    + \frac{\mq}{\Lambda} \left(\frac{h_1^q}{x}\right)^\star(\xbj) 
    + \frac{\mj-\mq}{\Lambda} \frac{h_1^q(\xbj)}{\xbj} 
\Biggr) \ ,
\end{align}
where we used $f^*(x) = f(x) - \int_x^1\frac{dy}{y} f(y)$, and $\tilde g_T$ and $\hat g_T$ are pure twist-3 functions that only depend on projections of quark-gluon-quark correlator, and are explicitly defined in that reference. The first four terms coincide with the result obtained in the conventional handbag
approximation \cite{Accardi:2009au}, while the last is new. Note that even if
the relation is written for the sum over quark flavors weighted by their charge
squared, it is also valid flavor by flavor; in fact, the steps
leading to such a decomposition are formulated at the quark correlator level.

@@

It is important to estimate the size of the various contributions to the non Wandzura-Wilczek part of $g_2$. We define the shorthand notation
\begin{align}
g_2^{\rm tw3} & = \frac{1}{2}\,\sum_q e_q^2
    \biggl(
    \widetilde g_T^{q \star}(\xbj) 
    + \int_\xbj^1\frac{dy}{y} \widehat{g}_T^q(y) 
    \biggr) 
\nonumber \\ 
g_2^{\text{quark}} &= \frac{1}{2}\,\sum_q e_q^2 
 \frac{\mq}{\Lambda} \bigg( \frac{h_1^q}{x} \bigg)^\star\!\!(\xbj) \ ,
\\
g_2^{\text{jet}} &= \frac{1}{2}\,\sum_q e_q^2 
\frac{\mj-\mq}{\Lambda} \frac{h_1^q(\xbj)}{\xbj} \ .
\nonumber 
\end{align} 
These terms are compared in Figure~\ref{f:g2contrib} to the $g_2-g_2^{WW}$ function obtained in the very recent JAM15 fit of polarized DIS asymmetries
\cite{Sato:2016tuz}, that includes a large amount of precise data at large $\xbj$ and small $Q^2$ from Jefferson Lab, and simultaneously fits the higher-twist components of
$g_1$ and $g_2$ to the data.\footnote{Notice, however, that the JAM15 fit imposes the $\int dx g_2(x) =0$ Burkhardt--Cottingham sum rule, which, however, is broken by inclusion of jet correlators, as discussed in Section~\ref{sec:moments}.} For the ``pure twist-3'' contribution,
$g_2^{\rm tw3}$, {\it i.e.}, the contribution from quark-gluon-quark matrix
elements, we show a recent light-front model calculation by Braun et al.~\cite{Braun:2011aw} (for bag model calculations, see \cite{Jaffe:1990qh,Stratmann:1993aw}). To estimate the contributions
from quark ($g_2^{\text{quark}}$) and jet mass ($g_2^{\text{jet}}$) effects, that depend on chiral-odd quark-quark matrix elements, we use the recent Pavia15 fit of the
transversity distribution from Ref.~\cite{Radici:2015mwa}, which is comparable
also to other 
extractions~\cite{Anselmino:2013vqa,Kang:2015msa}. Furthermore, we choose the
values of the mass parameters to be $\mq=5$ MeV and $\mj = 100$ MeV. 
%\AAcom{Did we really use 100 MeV in teh plots? I seem to remember 50 MeV.}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8cm]{g2contrib.pdf}
\includegraphics[width=8cm]{g2contribN.pdf}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left)
and neutron (right) $g_2$ structure functions compared to the JAM15 fit of the
$g_2-g_2^{\text{WW}}$ function (solid black) \cite{Sato:2016tuz}. The quark and jet
contributions are shown with a dotted red and a dot-dashed green line
respectively, with uncertainty bands coming form the Pavia15 fit of the
transversity function \cite{Radici:2015mwa}. The uncertainty in the choice
$m_q=5$ MeV and $\mj=100$ MeV is not shown. The pure twist-3 contribution
calculated by Braun et al. \cite{Braun:2011aw} is shown as a dashed blue line
(no uncertainty estimate was provided in the original reference). 
}
\end{center}
\end{figure}

@@

It is interesting to consider the moments of the non Wandzura-Wilczek contribution to $g_2$,
\begin{align}
  d_N \equiv (N+1) \int_0^1 dx\,x^N \bigg( g_2(x) - g_2^{WW}(x) \bigg) \ .
\end{align}
For a generic function $f$, let us define it's $N$-th moment as $f[N]=\int_0^1 dx\, x^{N} f(x)$. It is then straightforward to verify that $f^*[N] = f[N]\,N/(N+1)$ and  
\begin{align}
  d_N & = (N+1) g_2[N] + N g_1[N] \\
  & = \frac12 \sum_q e_q^2 \bigg[ N \tilde g_T^q[N] + \hat g_T^q[N]  
    + \frac{(N+1) \mj- \mq}{\Lambda} h_1^q[N-1] \bigg] \ .
\end{align}

>>>>

\pdfoutput=1 % if your are submitting a pdflatex to arXIv 
             % (i.e. if you have images in pdf, eps, png or jpg format)

@@

%\date{\bf Thursday, June 6th, 2017}
%\date{\bf \today, \currenttime}

\begin{abstract}
We revisit the standard analysis of inclusive Deep Inelastic Scattering
off nucleons taking into  account the fact that on-shell quarks cannot be present in the final state, but they rather decay into hadrons --  a process that can be described in terms of suitable ``jet'' correlators. As a consequence, a spin-flip term
associated with the invariant mass of the produced hadrons is generated
non perturbatively and couples to the target's transversity distribution
function. In inclusive cross sections, this provides an hitherto neglected and
large contribution to the twist-3 part of the $g_2$ structure function, that
can explain the discrepancy between recent calculations and fits of this
quantity.  It also provides an extension of the Burkhardt--Cottingham sum rule, putting constraints on the small-$x$ behavior of the transversity function, as well as an extension of the Efremov--Teryaev--Leader sum rule, suggesting a novel way to measure the tensor charge of the proton.
\end{abstract}

@@

The transversity distribution is notoriously difficult to access because it is
a chiral-odd function and needs to be combined with a spin-flip mechanism to
appear in a scattering process~\cite{Jaffe:1996zw}. Usually, this spin flip is provided by another
nonperturbative distribution or fragmentation function, accessible in Drell-Yan or semi-inclusive Deep Inelastic Scattering (DIS)~\cite{Ralston:1979ys,Jaffe:1991kp,Jaffe:1993xb,Collins:1992kk}.
The only other known way to attain spin-flip terms in Quantum Electro-Dynamics and QCD is taking into account mass corrections. In fact, it is well known that the transversity distribution gives a contribution to the structure function $g_2$ in inclusive DIS (see, \eg, \cite{Accardi:2009au} and references therein), and in particular to the violation of the so-called Wandzura--Wilczek relation for $g_2$~\cite{Wandzura:1977qf}. However, this contribution is proportional to the current quark mass and can be expected to be negligibly small.

@@

\begin{figure*}[bt]
  \centering
  (a)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram0}
  \hfill
  (b)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram2}
  \hfill
  (c)\includegraphics[width=0.3\linewidth,valign=t]{jetdiagram1}
  \caption{Diagrams contributing to inclusive DIS scattering up to twist-3, including a jet correlator in the top part. Note the gluon attaches to both the nucleon and jet correlators. The Hermitian conjugates of diagrams (b) and (c), i.e., with gluons attaching to the right of the cut, are not shown.
  }
  \label{fig:handbags}
\end{figure*}

When inserting the jet correlator in the handbag diagram for inclusive DIS, the integration over $dl^+$, or equivalently $dl^2/(2l^-)$, is kinematically coupled to the other integrations, and induces corrections of order ${\cal O}(1/Q^2)$ whose effect on the $F_2$ structure function has been studied in Ref.~\cite{Accardi:2008ne}. In this paper, where we limit our
attention to effects of order ${\cal O}(1/Q)$, we can neglect $k^-$ compared to $q^-$. As a consequence, we can extend the integration over $dl^2$ to infinity, with the consequence that the jet correlator decouples from the parton correlator $\Phi$, and the inclusive structure functions only depend on the integrated jet correlator 
\begin{equation} 
  \Xi(l^-,{\bm l_T}) \equiv \int \frac{dl^2}{2l^-} \, \Xi(l) 
    =  \frac{\Lambda}{2 l^-}\,\xi_1 \id
    +  \xi_2 \frac{\nslash_-}{2} 
    + {\cal O}(l_T/l^-) + \text{higher\ twists} \ .
\end{equation} 
The neglected ${\bm l_T}$-dependent and higher twist terms only contribute to ${\cal O}(1/Q^2)$ to the inclusive cross section. Note that thanks to Eq.~\eqref{eq:jetfnsprops} we obtain
\begin{align}
\xi_1 &= \int d\mu^2 \frac{\mu}{\Lambda} J_1(\mu^2) 
       \equiv \frac{\mj}{\Lambda},
&
\xi_2 &= \int d\mu^2 J_2(\mu^2) = 1 \ ,
\end{align} 
where $\mj$ can be interpreted as the average invariant mass produced in the spin-flip fragmentation processes of a quark of flavor $q$.
It is important to notice that while $\xi_2=1$ exactly due to CPT invariance
(see Sec.~10.7 of Ref.~\cite{Weinberg:1995mt}), the jet mass $0 < \mj < \int d\mu^2 \mu J_2(\mu^2)$ is dynamically determined. From the analytic properties of the spectral functions we know that $J_2(\mu^2) = Z \delta(\mu^2-m_q) + \bar J_2(\mu^2) \theta (\mu^2-m_\pi^2)$ with $Z<1$ and the continuum starting at $m_\pi$ (the mass of the pion) due to color confinement effects. Taking into account that $J_1 < J_2$, we may therefore estimate 
\begin{align}
  \label{eq:mjet}
  \mj = {\cal O}(100 \text{\ MeV}) \ .
\end{align}
Although $\mj$ is in general a nonperturbative quantity, it is interesting to
notice that on the perturbative vacuum, 
\begin{align*}
  \Xi^\text{pert}(l) = (\lslash + \mq \id) \, \delta(l^2-m_q^2) + {\cal O}(\alpha_s) \ ,
\end{align*}
where $\mq$ is the current quark mass; therefore $\mj^\text{pert}=m_q$, and one recovers the result of the calculation with the conventional handbag diagram.
However, we are here considering non perturbative effects
in the quark propagation, and $\mj \gg \mq$. 
Therefore, differently from $J_2$, the $J_1$ function leaves an imprint on the inclusive DIS cross section even in the asymptotic $Q^2 \to \infty$ regime.

@@

The role of the $\xi_1=M_q/\Lambda$ term in inclusive DIS can be discussed by analyzing the following terms of the semi-inclusive hadronic tensor \cite{Mulders:1995dh}:
\begin{align}
  \label{eq:Wsidis_ini}
  2 M  W^{\mu\nu}
    & = i \frac{2M}{Q} \hat t^{[\mu}_{\phantom \perp} 
    \epsilon_\perp^{\nu]\rho}S_{\perp\rho} \\
    & \times \sum_q e_q^2
    \bigg[ 2 \xbj g_T^q(\xbj) \sum_h \int d^2p_{hT} dz \, z\, D_1^{q,h}(z,p_{hT}) 
  + \frac{2\Lambda}{M} h_1^q(\xbj) \sum_h \int d^2p_{hT}  dz \, \tilde E^{q,h}(z,p_{hT}) \bigg] + \ldots \ , \nonumber
\end{align}
where $g_T^q(z,p_{hT})$ and $\tilde E^q(z,p_{hT})$ are twist-3 TMDs originating, respectively, from the quark-quark and the quark-gluon-quark fragmentation correlators. Note that in Eq.~\eqref{eq:Wsidis_ini} $M$ is the proton's mass, and we identified the power counting scale $M_h$ of Ref.~\cite{Mulders:1995dh} with our $\Lambda$. For clarity, we also reintroduced the quark flavors $q$, $e_q$ being their respective electric charge. 
The first term can be easily integrated with the help of the sum rules
\eqref{eq:SIDIS_to_DIS_TMDlevel1} and \eqref{eq:SIDIS_to_DIS_TMDlevel2}. To integrate the second term, however, we first need make
use of the relation $\tilde E(z) = E(z) - (\mq/\Lambda) z D_1(z)$, which is a
consequence of the QCD equations of motion \cite{Bacchetta:2006tn}, then
make again use of the sum rules \eqref{eq:SIDIS_to_DIS_TMDlevel1}-\eqref{eq:SIDIS_to_DIS_TMDlevel2} to obtain
\begin{align}
  \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) 
    = \xi_1 - \frac{\mq}{\Lambda} \xi_2 = \frac{\mj - \mq}{\Lambda }\ .
\end{align}
This formula provides us with a
non perturbative generalization of the commonly used $\int\tilde E =0$ sum
rule introduced in \cite{Jaffe:1996zw}. Indeed, calculating the jet correlator 
on the perturbative vacuum one would obtain, as already discussed, $\mj=\mq$
and the integral would vanish.

Finally, the contraction of the hadronic tensor with the leptonic tensor leads
to the following result for the inclusive DIS cross section up to order $M/Q$~\cite{Bacchetta:2006tn}:
\begin{align}
\frac{d\sigma}{d\xbj \, dy\, d\phi_S}
=
\frac{2 \alpha^2}{\xbj y Q^2}\,
\frac{y^2}{2\,(1-\varepsilon)}\, 
\biggl\{
&F_{T} + \varepsilon F_{L}
+ S_\parallel \lambda_e\,
  \sqrt{1-\varepsilon^2}\; 
F_{LL}
+ |\bm{S}_\perp| \lambda_e\, \sqrt{2\,\varepsilon (1-\varepsilon)}\, 
  \cos\phi_S\, 
F_{LT}^{\cos \phi_S}
 \biggr\} \ ,
\label{e:crossdis}
\end{align}
where $\phi_S$ is the angle between the transverse component of the proton spin vector and the lepton plane, $\epsilon$ is the ratio of the longitudinal and transverse photon fluxes, and $\lambda_e$ is the electron's helicity. 
The structure functions on the right hand side read
\begin{align}
F_{T} &= \xbj\,\sum_q e_q^2\,f_1^q(\xbj),
\\
F_{L} &= 0,
\\
F_{LL} &=\xbj\,\sum_q e_q^2\,g_1^q(\xbj),
\label{e:FLLint}
\\
F_{UT}^{\sin \phi_S}&=0,
\label{e:FUTint}
\\
F_{LT}^{\cos \phi_S}&=-\xbj\,\sum_q e_q^2\, \frac{2M}{Q}\,
\biggl(\xbj  g_T^q(\xbj)
   + \frac{\mj -\mq}{M} \, h_{1}^q(\xbj) \biggr) \ ,
\label{e:FLTint}
\end{align}
where $f_1^q$, $g_1^q$ and $h_1^q$ are the unpolarized, polarized, and transversity PDFs,respectively.
The second term in the last structure function is a new result from our
analysis, it is proportional to the jet mass, and it is not suppressed as an
inverse power of $Q$ compared to the 
standard $g_T$ term. Perturbatively, $M_q^\text{pert}=m_q$ and the new term vanishes. However, on the nonperturbative vacuum the jet mass $M_q$ is much larger than the quark's current mass $m_q$, originating a nonnegligible term to the twist-3 part of the target's $g_2$ structure function, as we will discuss in the next section.

@@

The structure functions in Eqs.~\eqref{e:FLLint} -- \eqref{e:FLTint} can be
related to the usual structure functions $g_1$ and $g_2$ defined from the following Lorentz decomposition of the antysymmetric part of the inclusive hadronic tensor
\begin{align}
\begin{split}
W_A^{\mu\nu}(P,q) 
  & = \frac{1}{P\cdot q} \varepsilon^{\mu\nu\rho\sigma} q_\rho 
  \Big[ S_\sigma g_1(x_B,Q^2) 
    + \Big( S_\sigma - \frac{S\cdot q}{P\cdot q}\, p_\sigma
	\Big) g_2(x_B,Q^2)
  \Big] \ .
\label{eq:Wmunu}
\end{split}
\end{align}
Then, neglecting contrinutions of order $1/Q^2$, one obtains \cite{Bacchetta:2006tn},
\begin{align}
  g_1 &= \frac{1}{2x_B} F_{LL} \\
  g_2+g_1 &= - \frac{Q}{4 x_B^2 M} F_{LT}^{\cos \phi_S} \ .
\end{align}
Utilizing equations of motion and Lorentz invariance relations as discussed in Ref.~\cite{Accardi:2009au} to decompose $g_T$ into ``pure twist-3'' and twist-2 pieces, we arrive at
\begin{align}
\label{e:g2}
  g_2(\xbj) = g_2^{WW}(\xbj) + \frac{1}{2}\,\sum_a e_a^2
\biggl(
    \widetilde g_T^{a \star}(\xbj) 
    + \int_\xbj^1\frac{dy}{y} \widehat{g}_T^q(y) 
    + \frac{\mq}{M} \left(\frac{h_1^q}{x}\right)^\star(\xbj) 
    + \frac{\mj-\mq}{M} \frac{h_1^q(\xbj)}{\xbj} 
\Biggr) \ ,
\end{align}
where we used $f^*(x) = f(x) - \int_x^1\frac{dy}{y} f(y)$, and $\tilde g_T$ and $\hat g_T$ are pure twist-3 functions that only depend on projections of quark-gluon-quark correlator, and are explicitly defined in that reference. The first four terms coincide with the result obtained in the conventional handbag
approximation \cite{Accardi:2009au}, while the last is new. Note that even if
the relation is written for the sum over quark flavors weighted by their charge
squared, it is also valid flavor by flavor; in fact, the steps
leading to such a decomposition are formulated at the quark correlator level.

@@

It is important to estimate the size of the various contributions to the non Wandzura-Wilczek part of $g_2$. We define the shorthand notation
\begin{align}
g_2^{\rm tw3} & = \frac{1}{2}\,\sum_q e_q^2
    \biggl(
    \widetilde g_T^{q \star}(\xbj) 
    + \int_\xbj^1\frac{dy}{y} \widehat{g}_T^q(y) 
    \biggr) 
\nonumber \\ 
g_2^{\text{quark}} &= \frac{1}{2}\,\sum_q e_q^2 
 \frac{\mq}{M} \bigg( \frac{h_1^q}{x} \bigg)^\star\!\!(\xbj) \ ,
\\
g_2^{\text{jet}} &= \frac{1}{2}\,\sum_q e_q^2 
\frac{\mj-\mq}{M} \frac{h_1^q(\xbj)}{\xbj} \ .
\nonumber 
\end{align} 
These terms are compared in Figure~\ref{f:g2contrib} to the $g_2-g_2^{WW}$ function obtained in the very recent JAM15 fit of polarized DIS asymmetries
\cite{Sato:2016tuz}, that includes a large amount of precise data at large $\xbj$ and small $Q^2$ from Jefferson Lab, and simultaneously fits the higher-twist components of
$g_1$ and $g_2$ to the data.\footnote{Notice, however, that the JAM15 fit imposes the $\int dx g_2(x) =0$ Burkhardt--Cottingham sum rule, which, however, is broken by inclusion of jet correlators, as discussed in Section~\ref{sec:moments}.} For the ``pure twist-3'' contribution,
$g_2^{\rm tw3}$, {\it i.e.}, the contribution from quark-gluon-quark matrix
elements, we show a recent light-front model calculation by Braun et al.~\cite{Braun:2011aw} (for bag model calculations, see \cite{Jaffe:1990qh,Stratmann:1993aw}). To estimate the contributions
from quark ($g_2^{\text{quark}}$) and jet mass ($g_2^{\text{jet}}$) effects, that depend on chiral-odd quark-quark matrix elements, we use the recent Pavia15 fit of the
transversity distribution from Ref.~\cite{Radici:2015mwa}, which is comparable
also to other 
extractions~\cite{Anselmino:2013vqa,Kang:2015msa}. Furthermore, we choose the
values of the mass parameters to be $\mq=5$ MeV and $\mj = 100$ MeV.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8cm]{g2contrib}
\includegraphics[width=8cm]{g2contribN}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left)
and neutron (right) $g_2$ structure functions compared to the JAM15 fit of the
$g_2-g_2^{\text{WW}}$ function (solid black) \cite{Sato:2016tuz}. The quark and jet
contributions are shown with a dotted red and a dot-dashed green line
respectively, with uncertainty bands coming form the Pavia15 fit of the
transversity function \cite{Radici:2015mwa}. The uncertainty in the choice
$m_q=5$ MeV and $\mj=100$ MeV is not shown. The pure twist-3 contribution
calculated by Braun et al. \cite{Braun:2011aw} is shown as a dashed blue line
(no uncertainty estimate was provided in the original reference). 
}
\end{center}
\end{figure}

@@

It is interesting to consider the moments of the non Wandzura-Wilczek contribution to $g_2$,
\begin{align}
  d_N \equiv (N+1) \int_0^1 dx\,x^N \bigg( g_2(x) - g_2^{WW}(x) \bigg) \ .
\end{align}
For a generic function $f$, let us define it's $N$-th moment as $f[N]=\int_0^1 dx\, x^{N} f(x)$. It is then straightforward to verify that $f^*[N] = f[N]\,N/(N+1)$ and  
\begin{align}
  d_N & = (N+1) g_2[N] + N g_1[N] \\
  & = \frac12 \sum_q e_q^2 \bigg[ N \tilde g_T^q[N] + \hat g_T^q[N]  
    + \frac{(N+1) \mj- \mq}{M} h_1^q[N-1] \bigg] \ .
\end{align}

The zero-th moment, $d_0=\int g_2$, provides an interesting relationship between transversity and the inclusive structure function $g_2$:
\begin{align}
  \label{eq:BC}
  \int dx\, g_2(x) = \sum_q e_q^2 \frac{\mj-\mq}{M} \int dx\, \frac{1}{x} h_1^q(x) \ .
\end{align}
Assuming $M_\text{light}\equiv M_u -m_u \approx M_d -m_d$ and dominance of light quarks in the sum over flavors, we can also write
\begin{align}
  \label{eq:BC-sfn}
  \int dx\, g_2(x) = \frac{M_\text{light}}{M} \int dx\, \frac{h_1(x)}{x} \ ,
\end{align}
with $h_1$ the transversity structure function.
